---
title: "Untitled"
format: html
editor: visual
---

## Packeges

```{r setup}
#install.packages(c("rvest", "tidyverse", "fs"))
library(rvest)
library(tidyverse)
library(fs)


study_years <- 2010:2020      # duration for this project 
```

\

```{r}
# WORKED EXAMPLE (complete — you don't need to change this)

cache_dir <- "raw_html"
dir_create(cache_dir)

get_page <- function(url, delay = 3) {
  # turn the URL into a safe filename
  file_name <- url %>%
    str_remove("^https?://") %>%
    str_replace_all("[/:?=&]", "_")
  file_path <- path(cache_dir, file_name)
  
  # if we already have it, read from disk (no request to the server)
  if (file_exists(file_path)) {
    return(read_html(file_path))
  }
  
  # otherwise: wait politely, download, save, return
  Sys.sleep(delay)
  message("Downloading: ", url)
  html <- read_html(url)
  xml2::write_html(html, file_path)
  html
}
```

# Project: Japanese Player Pages (2010–2020)

Test player from Taiwang

## Part 1: One Player Page

### Step 1: Read the page

```{r}
player_url  <- "https://npb.jp/bis/players/23525138.html"   # 王 柏融
player_html <- get_page(player_url)
player_html
```

```{r}
player_html |> html_element("title") |> html_text2()
```

### Step 2: Player ID

```{r}
player_id <- str_extract(player_url, "\\d+(?=\\.html)") # first \ is itself an escape character
player_id
```

### STEP 3: GET name and team

```{r}
page_title <- player_html %>%
  html_element("title") %>% # It finds the `<title>` from html
  html_text2() # remove the html tag like '<title>'

player_name <- str_extract(page_title, "^[^（]+")          # everything before （
player_team <- str_extract(page_title, "(?<=（)[^）]+")     # everything between （ and ）

player_name
player_team
```

### Step 4 : Find tables

```{r}
all_tables <- player_html %>% html_elements("table")
length(all_tables) # Theare are two tables

# peek at the first few rows of each one
all_tables %>%
  html_table(header = FALSE) %>%
  map(head, 3)
```

### Step 5: Get the bio table

```{r}
bio <- player_html %>%
  html_element("#pc_bio table") %>%    # the <table> inside the section with id pc_bio
  html_table(header = FALSE) |> 
  select(field = X1, value = X2) %>% # rename? 
  pivot_wider(names_from = field, values_from = value) %>% # reshape the tbales long to wide
  mutate(player_id = player_id,
         name = player_name,
         team = player_team,
         .before = 1) |> 
select(-any_of("身長／体重"))      

bio
```

```{r}
batting <- player_html %>%
  html_element("#pc_stats_wrapper #tablefix_b") %>% # #stats_b works as well
  html_table(header = TRUE) %>%
  filter(str_detect(年度,"^\\d{4}$")) %>%        # drop the career-totals row
  mutate(player_id = player_id, .before = 1)

batting
```

### Combine all

```{r}
scrape_player_jp <- function(url) {
  html <- get_page(url) #Part1 Step1
  # TODO
  player_id <- str_extract(url, "\\d{8}(?=\\.html)")       # Part 1, Step 2
  page_title <- html |> html_element("title") |> html_text2()       # Step 3
  
  bio <- html |>                                                     # Step 5
    html_element("#pc_bio table") |>
    html_table(header = FALSE) |>
    select(field = X1, value = X2) |>
    pivot_wider(names_from = field, values_from = value) |>
    mutate(player_id = player_id,
           name      = str_extract(page_title, "^[^（]+"),
           name_kana = html |> html_element("li#pc_v_kana") |> html_text2(),
           last_team = str_extract(page_title, "(?<=（)[^）]+"),
           .before = 1) |>
    select(-any_of("身長／体重"))
  
  # ---- batting ----
  batting_tbl <- html |> html_element("#pc_stats_wrapper #tablefix_b")
  if (inherits(batting_tbl, "xml_missing")) {
    batting <- NULL
  } else {
    batting <- batting_tbl |>
      html_table(header = TRUE) |>
      as_tibble(.name_repair = "unique") |>        # fixes empty column names
      filter(str_detect(年度,"^\\d{4}$")) %>%        # drop the career-totals row
      mutate(across(everything(), as.character)) |>
      mutate(player_id = player_id, .before = 1)
  }
  
  # ---- pitching ----
  pitching_tbl <- html |> html_element("#pc_stats_wrapper #tablefix_p")
  if (inherits(pitching_tbl, "xml_missing")) {
    pitching <- NULL
  } else {
    headers <- pitching_tbl |>
      html_elements(xpath = "./thead/tr/th") |>     # header
      html_text2()
    
    pitching <- pitching_tbl |>
      html_elements(xpath = "./tbody/tr") |>        # just only tbody, not tfoot, and Totals
      map(\(row) row |>
            html_elements(xpath = "./td") |>        # just td
            html_text2() |>
            as.list() |>
            set_names(headers) |>                   # stop when doens't match
            as_tibble_row()) |>
      list_rbind() |>
      mutate(across(everything(), \(x) str_remove_all(x, "\\s"))) |>   # remove the all space
      filter(str_detect(年度, "^\\d{4}$")) |>                          # remove 通算
      mutate(player_id = player_id, .before = 1)
  }                                                         
  # ---- stats (Part 1, Step 6), for each table that exists ----
  
  list(bio = bio, batting = batting, pitching = pitching)  
}                                                          


# TODO: Some pages will fail (typos, missing tables, a new player
#       with no stats yet). Wrap your function so ONE failure
#       doesn't kill the whole loop. Look up purrr::possibly().

safe_scrape_player <- possibly(scrape_player_jp, otherwise = NULL)

```

#### Test

```{r}
keppel <- scrape_player_jp("https://npb.jp/bis/players/41645131.html")
keppel$bio
keppel$pitching |> select(年度, 所属球団, 登板, 勝利, 投球回, 防御率)
wang <- scrape_player_jp("https://npb.jp/bis/players/23525138.html")
wang$batting#
```

## Part 2 Make it one roster page

Try Nippon_Ham 2012

```{r p-2 modify ver.}

roster_url <- "https://npb.jp/bis/players/search/yearly/2015/2005001"   # Nippon-Ham 2015
roster_html <- get_page(roster_url)

links <- roster_html %>% html_elements("a")

roster <- tibble(
  link_text = html_text2(links),
  href      = html_attr(links, "href")
) %>%
  filter(str_detect(href, "/players/\\d{8}\\.html")) %>%
  mutate(
    url        = url_absolute(href, roster_url),
    player_id  = str_extract(href, "\\d{8}"),
    name_jp    = str_split_i(link_text, "\n", 1) %>% str_trim(),
    debut_note = str_split_i(link_text, "\n", 2) %>% str_trim(),
    is_foreign_guess = str_detect(name_jp, "^[Ａ-Ｚ]．")　# this is not a final check
  ) %>%
  distinct(player_id, .keep_all = TRUE)

roster
nrow(roster)
```

Look correctly worked ( not include foreign players written in Kanji)

```{r}
roster %>%
  filter(is_foreign_guess) %>%
  select(name_jp, player_id, debut_note)
```

```{r}
player_links <- roster_html %>% html_elements(".player_unit_1") # old_player
length(player_links)
```

```{r}
roster_clean <- tibble(
  href = html_attr(player_links, "href"),
  name = player_links |> html_element(".name") |> html_text2(),
  note = player_links |> html_element(".note") |> html_text2() 
) |>
  mutate(
    url           = url_absolute(href, roster_url),       # full URL 
    player_id     = str_extract(href, "\\d{8}"),           # join key
  ) |>
  select(player_id, name, url)

roster_clean
```

## Part 3

### Step 1

```{r}
# コードが2010〜2020年で変わらない11球団
# 注意：constant_codes は2010年以降の球団コード。
#       2010年より前に広げる場合は、ダイエー・近鉄・旧ヤクルト・旧西武などのコードが必要。

constant_codes <- c(
  "1947001",   # 読売ジャイアンツ
  "1954001",   # 中日ドラゴンズ
  "1961001",   # 阪神タイガース
  "1968001",   # 広島東洋カープ
  "1992001",   # 千葉ロッテマリーンズ
  "2004001",   # 北海道日本ハムファイターズ
  "2005001",   # 福岡ソフトバンクホークス
  "2005002",   # オリックス・バファローズ
  "2005003",   # 東北楽天ゴールデンイーグルス
  "2006001",   # 東京ヤクルトスワローズ
  "2008001"    # 埼玉西武ライオンズ
)

constant_grid <- expand_grid(
  code = constant_codes,
  year = study_years 
)

# 横浜：2011年まで横浜ベイスターズ、2012年から横浜DeNAベイスターズ
yokohama_grid <- tibble(
  year = study_years,
  code = if_else(year <= 2011, "1993001", "2012001")
)

roster_grid <- bind_rows(constant_grid, yokohama_grid) |>
  mutate(url = str_glue("https://npb.jp/bis/players/search/yearly/{year}/{code}/"))

nrow(roster_grid) == 12 * length(study_years)   # True 
roster_grid |> filter(code %in% c("1993001", "2012001"))   # 横浜の11行を確認
```

```{r}
safe_scrape_roster_jp <- possibly(scrape_roster_jp, otherwise = NULL)

roster_results <- map(roster_grid$url, safe_scrape_roster_jp, .progress = TRUE)

all_rosters_jp <- roster_results |> list_rbind()
```

```{r}

# TODO: Turn your Part 2 code into a function, scrape_roster(url),
#       that returns a tibble of player links for ONE roster page.
scrape_roster_jp <- function(url) {
  html <- get_page(url)
  # TODO: your Part 2 code here
  
  # Create player_links 
  player_links <- html %>% html_elements(".player_unit_1")

  # roster clean (Part 2 Code)
  tibble(
    href = html_attr(player_links, "href"),
    name = player_links |> html_element(".name") |> html_text2(),
    note = player_links |> html_element(".note") |> html_text2()
  ) |>
    mutate(
      player_url         = url_absolute(href, url),            # url ではなく player_url
      player_id          = str_extract(href, "\\d{8}"),
      name_looks_foreign = str_detect(name, "^[Ａ-Ｚ]．"),
      year               = str_extract(url, "(?<=yearly/)\\d{4}"),
      team_code          = str_extract(url, "\\d{7}(?=/?$)")
    ) |>
    distinct(player_id, .keep_all = TRUE) |>
    select(player_id, name, note, name_looks_foreign, year, team_code, player_url)
}

# TODO: run it over all 12 rosters and stack the results
# hint: map() then list_rbind()   (or map_dfr())
# all_rosters <- ...

safe_scrape_roster_jp <- possibly(scrape_roster_jp, otherwise = NULL)

roster_results <- map(roster_grid$url, safe_scrape_roster_jp, .progress = TRUE)

all_rosters_jp <- roster_results |> list_rbind()
```

```{r}
roster_grid$url[map_lgl(roster_results, is.null)]       # 失敗したURL（空ならOK）
all_rosters_jp |> count(year, team_code) |> arrange(n)  # 132行、少ない順
nrow(all_rosters_jp)
n_distinct(all_rosters_jp$player_id)
```

```{r}
all_rosters_jp |> 
  filter(name == '李　大浩')
```

### Step 3: Tets

```{r}
players_jp <- all_rosters_jp |>
  distinct(player_id, .keep_all = TRUE)       # 1人1行（1,926行）

test_ids <- c(
  "51455119",   # 李 承燁   ：foreign player with kanji (b)
  "91695131",   # エドガー  ：no initial foreign player）(b)
  "23525114",   # W．オビスポ：foreign player with initial (p)
  "21125113",   # 阿部 慎之助：Japanese (b)
  "31135115"    # 朝井 秀樹 ：Japanese (p)
)

test_urls_jp <- players_jp |> filter(player_id %in% test_ids) |> pull(player_url)

test_results_jp <- map(test_urls_jp, safe_scrape_player_jp)

test_bio_jp      <- test_results_jp |> map("bio")      |> list_rbind()
test_batting_jp  <- test_results_jp |> map("batting")  |> list_rbind()
test_pitching_jp <- test_results_jp |> map("pitching") |> list_rbind()

test_urls_jp[map_lgl(test_results_jp, is.null)]               # 失敗（空ならOK）
test_bio_jp |> select(player_id, name, 経歴, ドラフト)         # 5行
test_batting_jp  |> filter(!str_detect(年度, "^\\d{4}$"))      # 0行
test_pitching_jp |> filter(!str_detect(年度, "^\\d{4}$"))      # 0行
```

## Run all the player (1926 players)

```{r}
player_results_jp <- map(players_jp$player_url, safe_scrape_player_jp, .progress = TRUE)

bio_jp      <- player_results_jp |> map("bio")      |> list_rbind()
batting_jp  <- player_results_jp |> map("batting")  |> list_rbind()
pitching_jp <- player_results_jp |> map("pitching") |> list_rbind()

failed_jp <- players_jp$player_url[map_lgl(player_results_jp, is.null)]
length(failed_jp)
```

### Save in csv

```{r}
write_csv(all_rosters_jp, "data/npb_rosters_2010_2020.csv")
write_csv(bio_jp,         "data/npb_bio_jp.csv")
write_csv(batting_jp,     "data/npb_batting_jp.csv")
write_csv(pitching_jp,    "data/npb_pitching_jp.csv")
```

```{r}
nrow(bio_jp)
```

### Check foreign player of not

```{r}

```

#### 43 players: name -\> `false` and draft -\> `true`

```{r}
review <- foreign_check |>
  filter(name_looks_foreign != draft_empty) |>
  select(player_id, name, 経歴, ドラフト, name_looks_foreign, draft_empty) |>
  arrange(name_looks_foreign)

review |> print(n = Inf)
```

```{r}
player_url <- "https://npb.jp/bis/players/63165116.html"   # T. Sledge
player_html <- get_page(player_url)

# --- what's inside the name block? ---
as.character(player_html %>% html_element("div#pc_v_name"))

# --- name: Japanese, kana, and English ---
name_jp   <- player_html %>% html_element("li#pc_v_name") %>% html_text2() %>% str_trim()
name_kana <- player_html %>% html_element("li#pc_v_kana") %>% html_text2() %>% str_trim()

name_en <- name_kana %>%
  str_extract("\\(([^)]+)\\)") %>%
  str_remove_all("[()]") %>%
  str_trim()

name_jp     # "Ｔ．スレッジ"
name_kana   # "ターメル・スレッジ (TERRMEL SLEDGE)"
name_en     # "TERRMEL SLEDGE"

# --- is this a batter or a pitcher? ---
has_batting  <- !is.na(player_html %>% html_element("#tablefix_b"))
has_pitching <- !is.na(player_html %>% html_element("#tablefix_p"))

has_batting
has_pitching

# --- bio block ---
as.character(player_html %>% html_element("section#pc_bio"))

# --- batting stats, with year and team ---
bat_raw <- player_html %>% html_element("#tablefix_b") %>% html_table()
bat_raw

# year and team have their own classes, so pull them directly if the
# html_table() version comes out messy
rows <- player_html %>% html_elements("#tablefix_b tr.registerStats")

bat_keys <- tibble(
  year = rows %>% html_element("td.year") %>% html_text2() %>% str_trim(),
  team = rows %>% html_element("td.team") %>% html_text2() %>% str_trim()
)

bat_keys
```

```{r}
# 成績表を読む（入れ子の表の問題に対応）
parse_stats_table <- function(tbl, player_id) {
  if (inherits(tbl, "xml_missing")) return(NULL)
  
  headers <- tbl |> html_elements(xpath = "./thead/tr/th") |> html_text2()
  
  tbl |>
    html_elements(xpath = "./tbody/tr") |>
    map(\(row) row |>
          html_elements(xpath = "./td") |>
          html_text2() |>
          as.list() |>
          set_names(headers) |>
          as_tibble_row()) |>
    list_rbind() |>
    filter(str_detect(年度, "^\\d{4}$")) |>        # remove 通算
    mutate(across(everything(), \(x) str_remove_all(x, "\\s"))) |>
    mutate(player_id = player_id, .before = 1)
}

scrape_player_jp <- function(url) {
  html      <- get_page(url)
  player_id <- str_extract(url, "\\d{8}(?=\\.html)")
  
  bio <- html |>
    html_element("#pc_bio table") |>
    html_table(header = FALSE) |>
    select(field = X1, value = X2) |>
    pivot_wider(names_from = field, values_from = value) |>
    mutate(player_id = player_id,
           name_jp   = html |> html_element("li#pc_v_name") |> html_text2(),
           name_kana = html |> html_element("li#pc_v_kana") |> html_text2(),
           .before = 1) |>
    select(-any_of("身長／体重"))                     
  
  list(
    bio      = bio,
    batting  = parse_stats_table(html_element(html, "#tablefix_b"), player_id),
    pitching = parse_stats_table(html_element(html, "#tablefix_p"), player_id)
  )
}
```

```{r}
akashi <- scrape_player_jp("https://npb.jp/bis/players/51155119.html")   # 日本人・ドラフトあり
lee    <- scrape_player_jp("https://npb.jp/bis/players/61865135.html")   # 韓国・ドラフト空欄
wang   <- scrape_player_jp("https://npb.jp/bis/players/23525138.html")   # 台湾・ドラフト空欄

akashi$bio; lee$bio; wang$bio
lee$batting
```

```{r}
foreign_check <- foreign_check |>
  mutate(is_foreign = draft_empty)

foreign_check |> count(is_foreign)   # 363 + 43 = 406 players
```

### Position and Arrival year

```{r}
# players with at least one pitching row
pitcher_ids <- unique(pitching_jp$player_id)
hitter_ids  <- unique(batting_jp$player_id)

# earliest NPB season, from both tables
first_season <- bind_rows(
  batting_jp  |> select(player_id, 年度),
  pitching_jp |> select(player_id, 年度)
) |>
  group_by(player_id) |>
  summarise(first_year = min(as.integer(年度)))

foreign_players <- foreign_check |>
  filter(is_foreign) |>
  mutate(position = case_when(
    player_id %in% pitcher_ids ~ "pitcher",
    player_id %in% hitter_ids  ~ "position",
    TRUE                       ~ "no_stats"
  )) |>
  left_join(first_season, by = "player_id")

foreign_players |> count(position)
```

```{r}
foreign_players |>
  count(position, in_window = between(first_year, 2010, 2020))
```

```{r}
study_players <- foreign_players |>
  filter(position == "position",
         between(first_year, min(study_years), max(study_years)))

nrow(study_players)    # 148

write_csv(study_players, "data/study_players.csv")
```

```{r}
study_players <- study_players |>
  mutate(prev_club = str_split_i(経歴, " - ", -1))

study_players |> count(prev_club, sort = TRUE) |> print(n = Inf)
```

MLB - 99 players

minor - 26 players

total 125 players.

```{r}
study_players <- study_players |>
  mutate(prev_league = case_when(
    str_detect(prev_club, "キューバ")    ~ "Cuba",
    str_detect(prev_club, "韓国")        ~ "KBO",
    str_detect(prev_club, "台湾")        ~ "CPBL",
    str_detect(prev_club, "^BC|四国")    ~ "Japan indie",
    str_detect(prev_club, "米独立")      ~ "US indie",
    str_detect(prev_club, "ドミニカ")    ~ "Dominican",
    TRUE                                 ~ "MLB/MiLB"
  ))

study_players |> count(prev_league, sort = TRUE)
```

For now analize 2010 - 2025

## Create the low English name and birth date

```{r}
study_players <- study_players |>
  mutate(
    name_en    = str_extract(name_kana, "(?<=[\\(（]).+(?=[\\)）])") |> str_trim(),
    birth_date = ymd(生年月日)
  )

study_players |> select(name, name_kana, name_en, 生年月日, birth_date) |> head(10)
study_players |> filter(is.na(name_en) | is.na(birth_date))   # the player has no english name
```

### Check the position (it can use to identify pitcher or batter) 

```{r}
bio_jp |> count(ポジション, sort = TRUE)
```

# Baseballr

```{r}
library(baseballr)

chadwick <- chadwick_player_lu()

names(chadwick)
nrow(chadwick)
```

Actually, they have `key_npb` ! (And `birth_date` as well)

## Try to make sure I can use `key_npb`

```{r}
chad_npb <- chadwick |>
  filter(!is.na(key_npb), key_npb != "") |>
  select(key_npb, key_mlbam, key_bbref_minors, name_first, name_last,
         birth_year, birth_month, birth_day)

nrow(chad_npb)          # the number of player has 'key_npb'
head(chad_npb, 100)          # make sure what kind of form
chad_npb |> filter(str_detect(name_last, "Keppel"))   # Using ケッペル（41645131）as an example
```

Yey! They actually have the same id `key_npb`
